DATA 621 - Business Analytics and Data Mining

### 2020-09-24 HW2/HW2-Group2.rmd
Finished the HW

# 9. Bounds on \(F_1\)
## Proof 1
We can prove this by decomposing the metrics into their component parts as
defined above.
\[
\begin{align}
F_{1:numerator} &= 2 \cdot P \cdot S\\
&= 2 \cdot \frac{TP}{TP + FP} \cdot \frac{TP}{TP + FN}\\
&= \frac{2\cdot(TP)^2}{(TP+FP)(TP+FN)}\\
F_{1:denominator} &= P + S\\
&= \frac{TP}{TP + FP} + \frac{TP}{TP + FN}\\
&= \frac{TP(TP + FN) + TP(TP+FP)}{(TP+FP)(TP+FN)}\\
&= \frac{(TP)^2 + TP\cdot FN + (TP)^2 + TP\cdot FP}{(TP+FP)(TP+FN)}\\
&= \frac{2\cdot(TP)^2 + TP\cdot (FN + FP)}{(TP+FP)(TP+FN)}\\\
F_1 &= \frac{F_{1:numerator}}{F_{1:denominator}}\\
&= \frac{2\cdot(TP)^2}{2\cdot(TP)^2 + TP\cdot (FN + FP)}\\
&= \frac{2\cdot TP}{2\cdot TP + FN + FP}
\end{align}
\]

@@

10. ROC Curve
The receiver operating curve (ROC) is defined as a plot of the true positive
rate (TPR) against the false positive rate (FPR) at different probability
thresholds. The FPR is usually plotted on the x-axis. The function below takes
as input a data frame with columns named `class` and `pred_class` as above, and
a third column named `pred_prob` which reflects the predicted probability of
being positive. It then steps along a sequence from 0 to 1, assigning classes
based on that sequence, calculating the TPR and FPR for each threshold, plotting
the results when completed. Lastly, it calculates the area under the ROC (AUC).
Trapezoidal quadrature will be used to estimate the AUC from the table created
for the ROC, as the intervals on the x axis are not equidistant.

```{r ROC}
ROC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    rocTable <- data.frame(thresh = seq(0, 1, 0.01),
                           x = double(101),
                           y = double(101))
    for (i in seq_along(rocTable$thresh)) {
      dftmp <- data.frame(class = df$class,
                          pred_class = ifelse(df$pred_prob > rocTable$thresh[i],
                                              1, 0))
      rocTable$x[i] <- 1 - SPC(dftmp)
      rocTable$y[i] <- SNS(dftmp)
    }
  }
  return(list(plot = {plot(rocTable$x, rocTable$y, type = 'l',
                          xlab = "False Positive Rate",
                          ylab = 'True Positive Rate')
    lines(c(0, 1), c(0, 1), lty = 2)},
              AUC = sum((rocTable$y[1:100] + rocTable$y[2:101]) * 0.5 *
                          (rocTable$x[1:100] - rocTable$x[2:101]))))
}
```

@@

```{r ROCdata, echo=FALSE, fig.height=6}
ROC(class_data)
```

@@

```{r sec11}
ACC(class_data)
CER(class_data)
PRC(class_data)
SNS(class_data)
SPC(class_data)
F1(class_data)
ROC(class_data)
```

@@

# 13.

>>>>

# 9. Bounds on \(F_1\)
## Proof 1
We can prove this by decomposing the metrics into their component parts as
defined above.
\[
\begin{aligned}
F_{1:numerator} &= 2 \cdot P \cdot S\\
&= 2 \cdot \frac{TP}{TP + FP} \cdot \frac{TP}{TP + FN}\\
&= \frac{2\cdot(TP)^2}{(TP+FP)(TP+FN)}\\
F_{1:denominator} &= P + S\\
&= \frac{TP}{TP + FP} + \frac{TP}{TP + FN}\\
&= \frac{TP(TP + FN) + TP(TP+FP)}{(TP+FP)(TP+FN)}\\
&= \frac{(TP)^2 + TP\cdot FN + (TP)^2 + TP\cdot FP}{(TP+FP)(TP+FN)}\\
&= \frac{2\cdot(TP)^2 + TP\cdot (FN + FP)}{(TP+FP)(TP+FN)}\\\
F_1 &= \frac{F_{1:numerator}}{F_{1:denominator}}\\
&= \frac{2\cdot(TP)^2}{2\cdot(TP)^2 + TP\cdot (FN + FP)}\\
&= \frac{2\cdot TP}{2\cdot TP + FN + FP}
\end{aligned}
\]

@@

# 10. ROC Curve
The receiver operating curve (ROC) is defined as a plot of the true positive
rate (TPR) against the false positive rate (FPR) at different probability
thresholds. The FPR is usually plotted on the x-axis. The function below takes
as input a data frame with columns named `class` and `pred_class` as above, and
a third column named `pred_prob` which reflects the predicted probability of
being positive. It then steps along a sequence from 0 to 1, assigning classes
based on that sequence, calculating the TPR and FPR for each threshold, plotting
the results when completed. Lastly, it calculates the area under the ROC (AUC).
Trapezoidal quadrature will be used to estimate the AUC from the table created
for the ROC, as the intervals on the x axis are not equidistant.

```{r ROC}
ROC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    rocTable <- data.frame(thresh = seq(0, 1, 0.001),
                           x = double(1001),
                           y = double(1001))
    for (i in seq_along(rocTable$thresh)) {
      dftmp <- data.frame(class = df$class,
                          pred_class = ifelse(df$pred_prob > rocTable$thresh[i],
                                              1, 0))
      rocTable$x[i] <- 1 - SPC(dftmp)
      rocTable$y[i] <- SNS(dftmp)
    }
  }
  return(list(plot = {plot(rocTable$x, rocTable$y, type = 'l',
                          xlab = "False Positive Rate",
                          ylab = 'True Positive Rate')
    lines(c(0, 1), c(0, 1), lty = 2)},
              AUC = sum((rocTable$y[1:1000] + rocTable$y[2:1001]) * 0.5 *
                          (rocTable$x[1:1000] - rocTable$x[2:1001]))))
}
```

@@

```{r ROCdata, echo=FALSE, fig.height=6, fig.width=6}
ROC(class_data)
```

@@

```{r sec11}
ACC(class_data)
CER(class_data)
PRC(class_data)
SNS(class_data)
SPC(class_data)
F1(class_data)
```

@@

# 13. Using the `pROC` package

```{r pROC, fig.height=6, fig.width=6}
ROC(class_data)
library(pROC)
pROCroc <- roc(class_data, class, pred_prob)
plot(pROCroc)
auc(pROCroc)
```

Visual inspection of the two ROC curves shows extremely similar shapes. The AUC
scores are also practically identical. Originally, we used the suggested 101
points between 0 and 1. This led to a difference of less than 1% in the AUC
scores. Moving to a 1001-point quadrature all but eliminated that difference.
